#Reproducible Research Course Project

##Loading and Preprocessing the Data

###Load the data
```{r}
readData <- read.csv("activity.csv")
```
###Let's summarize our data
```{r}
head(readData)
str(readData)
```
#What is mean total number of steps taken per day?

##1.Calculate the total number of steps taken per day.
```{r}
sumsteps <- aggregate(steps ~ date,data = readData,sum)
head(sumsteps)
```
##2.Make a histogram of the total number of steps taken each day.
```{r}
hist(sumsteps$steps,col=8,main = "Histogram of the total number of steps taken each day",
     xlab = "Total number of steps in a day",ylab = "Frequency")
```





##3.Calculate and report the mean and median of the total number of steps taken per day.

```{r}
mean(sumsteps$steps)  
median(sumsteps$steps)

```
#What is the average daily activity pattern?

##1.Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).

```{r}
averageSteps <- aggregate(steps ~ interval,data = readData,mean)

plot(averageSteps$interval,averageSteps$steps,type = "l",main="Average number of steps taken",
     xlab = "Interval",ylab = "steps")

```




##2.Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
maxStepInterval <- averageSteps[which.max(averageSteps$steps),]
maxStepInterval

```
#Imputing missing values

##1.Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NA's.
```{r}
sum(is.na(readData))

```
##2.Devise a strategy for filling in all of the missing values in the dataset.
###I am using the mean strategy to fill the missing values.
```{r}
imputeData <- readData
imputeData$steps[which(is.na(readData$steps))] <- mean(imputeData$steps,na.rm = TRUE)

```
##3.Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
newDataset <- as.data.frame(imputeData)
```


##4a.Make a histogram of the total number of steps taken each day.
###Here we have selected the data with no missing values.
```{r}
StepsTotal <- aggregate(steps ~ date, data = newDataset, sum, na.rm = TRUE)

hist(StepsTotal$steps, main = "Total number of steps taken each day", xlab = "day",col = 8)

```



##4b.Calculate and report the mean and median total number of steps taken per day.
```{r}
oldMean <- mean(sumsteps$steps)
oldMean
oldMedian <- median(sumsteps$steps)
oldMedian
```
###New Mean and Median
```{r}
newMean <- mean(StepsTotal$steps)
newMean
newMedian <- median(StepsTotal$steps)
newMedian
```
##4c.Do these values differ from the estimates from the first part of the assignment? 
###Yes, the mean is the same but the median is increased by 1.19 steps.


#5.Are there differences in activity patterns between weekdays and weekends?
##5a.Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
dateData <- as.POSIXct(newDataset$date)
newDataset$weekday <- ifelse(weekdays(dateData) %in% c("Saturday", "Sunday"), 
                           "weekend", "weekday")

```
##5b.Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
```{r}
intervalData <- aggregate(steps ~ interval + weekday,data = newDataset,mean)
library(ggplot2)
ggplot(intervalData, aes(x=interval, y=steps)) + geom_line()+
  facet_grid(weekday ~.) + xlab("Interval") + ylab("Steps") +
  ggtitle("Average Number of Steps in Each Interval")

```



